Toy Transit Surveys

### 2017-08-30 memos/workdir/toy_analytic_surveys.tex
model 2: fix presentation of correction factor

Astronomer A misclassifies planetary radii, miscounts the total number of 
stars, and (incorrectly) estimates his completeness and geometric correction 
for ``single'' stars as a function of radius $f(R_p')$ (omitting a ``s'' 
subscript because the stars for which this incorrect occurrence rate are 
derived are not all single).
His occurrence rate for planets of radius \textit{near} $R_p$, \textit{i.e.} 
the set of planets with $\{ (1-x)R_p < R_p^{\rm obs} < R_p \}$, is
\begin{equation}
\Gamma_{{\rm A}, R_p'} = \frac{ (N_{\rm det,s} + N_{\rm det,d}(R_p')) 
/ f_s(R_p')   }{ N_s + N_d}.
\end{equation}

For him,
\begin{equation}
f(R_p') = f_{g}(R_p') f_{c}(R_p').
\end{equation}

Astronomer D, who had everything right, doesn't change anything.

%TODO: LEFT OFF HERE!

The new difference between the two rates is
\begin{align}
\Delta' &\equiv \left|
\frac{\Gamma_{{\rm A},R_p'} - \Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} }
\right| \\
&= \left|
1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} } 
\right| \\
&=
\left| 1 -
\left( \frac{1+\chi \beta}{1 + \xi}  \cdot 
\frac{1 + \beta}{1 + 2\beta} \right)
\right|,
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s, \quad \chi \equiv (\Gamma_{t,d1} + 
\Gamma_{t,d2})/\Gamma_{t,s},
\end{equation}
and
\begin{equation}
\xi \equiv \frac{N_{\rm det,d}(R_p')}{f_s(R_p')} \cdot \frac{1}{N_s 
\Gamma_{t,s}}.
\end{equation}
If we let $x < 1 - 2^{-1/2}$, for instance $x=0.1$, then $N_{d2}(R_p') = 0$. 
The expression for $\xi$ simplifies to
\begin{equation}
\chi = \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
\frac{f_{d1}(R_p')}{f_s(R_p')},\quad ({\rm if\ }x < 1-2^{-1/2}).
\label{eq:chi_simplfied}
\end{equation}

If we assume $\Gamma_{t,d1} = \Gamma_{t,s}$, as we were doing 
anyway, Eq.~\ref{eq:chi_simplfied} is bounded above by
$\beta$. This is because $N_{d1}(R_p') \leq N_{d1} = N_d$, and $f_{d1}(R_p') / 
f_s(R_p') \leq 1$, since over a fixed radius interval dilution will always 
yield a lower completeness fraction for double star systems than singles.
Recall for the fixed $\gamma_R$ population we showed that $f_{d1}/f_s = 
(1+\gamma_R)^{-3}$.

Regardless, the cheap way of making the estimate without deriving the actual 
completeness fractions, or even running numerics, is just to plot the error as 
a function of $\xi$. We do this in Fig.~\ref{fig:deltaprime_vs_xi}, which 
shows that the
relative error varies between $\approx 5\% - 35\%$, depending on what 
fraction of secondaries you assume host planets.

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/deltaprime_vs_xi.pdf}
	\end{center}
	\caption{\textit{Y axis}: relative error between the Astronomer who ignores 
	binarity 
	but counts planetary radii that are ``close enough'' (A') and the Astronomer 
	who knows everything (D).
	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplfied}. Taking a 
	guess at the true value, if $N_{d1}(R_p') = N_d/2$, and the occurrence rates 
	between `d1' and `s' are the same, and we put $\gamma_R = 0.1$ into the 
	completeness ratio, then $\xi = (\beta/2) \times (1+0.1)^{-3} = 0.22$. The 
	relative error then varies between $\approx 5\% - 35\%$, depending on what 
	fraction of secondaries you assume host planets.
		}
	\label{fig:deltaprime_vs_xi}
\end{figure}

@@

We can attempt an answer with Monte Carlo simulations of something like the 
\kepler\ field. I say ``something like'' because the binarity properties of the 
\kepler\ field (or more specifically the stars selected in the KIC) are not 
known.
\todo[inline]{but it might be worth doing something directly on the KIC's 
selected target stars, like what Ciardi did! The following is a good example, 
but at most it can ``suggest'' that a similar level of error has been made from 
the \kepler\ survey.}

@@

Similar to Fig.~\ref{fig:deltaprime_vs_xi}, Fig.~\ref{fig:error_vs_occrate} 
shows that the magnitude of the relative error 
made by the $\Gamma_{{\rm A},R_p'}$ and $\Gamma_{{\rm A},R_p}$ estimates 
depends strongly 
on the fraction of secondaries that are assumed to host the $(R_p, P)$ planets.
The simplest assumption is that the occurrence rate is the same for all solar 
type stars: any ``Sun-like'' ($0.7<M_\star/M_\odot<1.3$) star hosts planets at 
the ``true rate'' $\Gamma_t$, and any other mass of star does not.
Given how we have defined the GIC, this means that all single stars and 
primaries of double star systems are sun-like. The fraction of Sun-like 
secondaries in this case is 37.4\% of all secondaries.
Reading the appropriate value off Fig.~\ref{fig:error_vs_occrate}, this 
corresponds to $\Gamma_{\rm D} = 1.39 \Gamma_{\rm A'}$.

>>>>

Astronomer A misclassifies planetary radii, miscounts the total number of 
stars, and (incorrectly) estimates his completeness and geometric correction 
for ``single'' stars as a function of radius $f(R_p')$ (omitting a ``s'' 
subscript because the stars for which this incorrect occurrence rate are 
derived are not all single).
His occurrence rate for planets of radius \textit{near} $R_p$, \textit{i.e.} 
the set of planets with $\{ (1-x)R_p < R_p^{\rm obs} < R_p \}$, is
\begin{equation}
\Gamma_{{\rm A}, R_p'} = \frac{ (N_{\rm det,s} + N_{\rm det,d}(R_p')) 
/ f(R_p')   }{ N_s + N_d}.
\end{equation}

For him,
\begin{equation}
f(R_p') = f_g(R_p') f_c(R_p').
\end{equation}

Astronomer D, who had everything right, changes nothing.

The new correction factor is
\begin{align}
X_\Gamma &= \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} } \\
&=
\frac{1+\chi \beta}{1 + \xi}  \cdot 
\frac{1 + \beta}{1 + 2\beta},
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s, \quad \chi \equiv (\Gamma_{t,d1} + 
\Gamma_{t,d2})/\Gamma_{t,s},
\end{equation}
and
\begin{equation}
\xi \equiv \frac{N_{\rm det,d}(R_p')}{f(R_p')} \cdot \frac{1}{N_s 
\Gamma_{t,s}}.
\end{equation}
If we let $x < 1 - 2^{-1/2}$, for instance $x=0.1$, then $N_{d2}(R_p') = 0$. 
The expression for $\xi$ in this case becomes
\begin{align}
\xi &= \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
\frac{f_{d1}(R_p')}{f(R_p')} \\
&= \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
\frac{f_{g,d1}(R_p')}{f_g(R_p')}\cdot \frac{f_{c,d1}(R_p')}{f_c(R_p')},
\label{eq:chi_simplified}
\end{align}
provided that $x < 1-2^{-1/2}$.

We have now parametrized our ignorance.
The values that the latter two ratios of Eq.~\ref{eq:chi_simplified} take 
depend on what we assume Astronomer A' does to derive the stellar parameters 
for the systems that he thinks are singles but are in fact doubles.
For instance, assume he somehow observed the bolometric flux from every point 
on the sky, and used it with a distance to obtain single-star luminosities, and 
thus radii and masses.
His binary systems (thought to be singles) would be bluer than either component 
star is in reality.
On the main sequence, $\rho_\star$ decreases with increasing stellar radius, 
so $\rho_\star$ would be underestimated for all binaries.
Since $f_g \propto \rho_\star^{1/3}$, this means that $f_{g,d1}(R_p')/f_g(R_p') 
> 1$.
However, by the same token we can place an \textit{upper} bound on this ratio. 
In the worst case scenario a given star's luminosity is twice the individual 
star's luminosity. If $L\propto M^3 \propto R^3$, then the radius is 
over-estimated by at most a factor of $2^{1/3}$. Thus the density is 
underestimated by at most a factor of $2^{1/2}$, and $f_g(R_p')$ is 
underestimated by at most a factor of $2^{1/6}=1.12$.

This begins to take us somewhere. The left-most term of 
Eq.~\ref{eq:chi_simplified} is bounded above by $\beta$. This is because 
$N_{d1}(R_p') \leq N_{d1} = N_d$.
If we assume $\Gamma_{t,d1} = \Gamma_{t,s}$, as we were doing 
anyway, then the product of the three left-most fractions must be less than 
$1.12\beta$.
The only thing standing in our way of an analytic upper bound to $\xi$ is the 
completeness fraction $f_{c,d1}(R_p')/f_c(R_p')$.
Note that over a fixed radius interval dilution will always 
yield a lower completeness fraction for double star systems than singles, so 
$f_{c,d1}(R_p') / f_{c,s}(R_p') \leq 1$.

For the completeness term, an argument that I don't wholly believe
is that
\begin{equation}
\frac{f_{c,d1}(R_p')}{f_c(R_p')} = \frac{f_{c,d1}(R_p')}{f_{c,s}(R_p')} \cdot
	\frac{f_{c,s}(R_p')}{f_{c}(R_p')} \lessapprox 1 \cdot (1+\langle \gamma_R 
	\rangle)^3,
\end{equation}
where the first part of the inequality is obvious, but the second is by 
assuming it's $\lessapprox f_{s,c}/f_{d,c}$, which might be wrong because of 
the radius dependence.

A way of making the estimate without deriving the actual 
completeness fractions, or even running numerics, is just to plot the error as 
a function of $\xi$. We do this in Fig.~\ref{fig:XGamma_vs_xi}, which 
shows that the correction factor varies from 0.6 to 1.6, depending on what 
fraction of secondaries you assume host planets.
\todo[inline]{I could run numerics for this to make the result more 
interesting.}

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/XGamma_vs_xi.pdf}
	\end{center}
	\caption{\textit{Y axis}: correction factor for the astronomer who 
	ignores binarity but counts planetary radii that are ``close enough'' (A') 
	and the astronomer who knows everything (D).
	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplified}. Taking a 
	guess at the true value, if $N_{d1}(R_p') = N_d/2$, and the occurrence rates 
	between `d1' and `s' are the same, and we put $\gamma_R = 0.1$ into the 
	completeness ratio, then $\xi = (\beta/2) \times (1+0.1)^{-3} = 0.22$. The 
	relative error then varies between $\approx 5\% - 35\%$, depending on what 
	fraction of secondaries you assume host planets.
		}
	\label{fig:XGamma_vs_xi}
\end{figure}

@@

We can attempt an answer with Monte Carlo simulations of something like the 
\kepler\ field. I say ``something like'' because the binarity properties of the 
\kepler\ field (or more specifically the stars selected in the KIC) are not 
known.
\todo[inline]{but it might be better to do something directly on the KIC's 
selected target stars, like what Ciardi did! The following is a good example, 
but at most it can ``suggest'' that a similar level of error has been made from 
the \kepler\ survey.}

@@

Similar to Fig.~\ref{fig:XGamma_vs_xi}, Fig.~\ref{fig:error_vs_occrate} 
shows that the magnitude of the relative error 
made by the $\Gamma_{{\rm A},R_p'}$ and $\Gamma_{{\rm A},R_p}$ estimates 
depends strongly 
on the fraction of secondaries that are assumed to host the $(R_p, P)$ planets.
The simplest assumption is that the occurrence rate is the same for all solar 
type stars: any ``Sun-like'' ($0.7<M_\star/M_\odot<1.3$) star hosts planets at 
the ``true rate'' $\Gamma_t$, and any other mass of star does not.
Given how we have defined the GIC, this means that all single stars and 
primaries of double star systems are sun-like. The fraction of Sun-like 
secondaries in this case is 37.4\% of all secondaries.
Reading the appropriate value off Fig.~\ref{fig:error_vs_occrate}, this 
corresponds to $\Gamma_{\rm D} = 1.39 \Gamma_{\rm A'}$.
